---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data <- read.csv('activity.csv')
```
## What is mean total number of steps taken per day?
### The total number of steps taken per day
```{r}
daily_steps <- tapply(data$steps, 
                      data$date, 
                      function(x){sum(x, na.rm = TRUE)})
head(daily_steps)
```
### Histogram of the total number of steps taken each day
```{r}
hist(daily_steps, main = 'Steps Per Day Histogram', xlab = '# steps per day')
```

### Mean of steps per day
```{r}
mean(daily_steps, na.rm = TRUE)
```
### Median of steps per day
```{r}
median(daily_steps, na.rm = TRUE)
```
## What is the average daily activity pattern?
```{r}
interval_steps <- tapply(data$steps, 
                      data$interval, 
                      function(x){mean(x, na.rm = TRUE)})
```
### Timeseries plot
```{r}
plot(y=interval_steps,
     x=names(interval_steps),
     type = 'l', 
     xlab = 'Interval', 
     ylab = 'Avarage # of steps')
```

### Interval with max number of steps
```{r}
sprintf("Interval with max number of steps: %s",
        names(interval_steps)[which.max(interval_steps)])
```

## Imputing missing values
### Number of rows with NA
```{r}
sum(apply(is.na(data),1,any))
```
### Fill NAs by mean of interval
```{r}
library(tidyr)
library(dplyr)
nan_mask = apply(is.na(data), 1, any)
filled_data <- data %>%
  group_by(interval) %>%
  mutate_all(funs(ifelse(is.na(.), mean(., na.rm = TRUE),.)))
```
### Histogram of the total number of steps taken each day (filled NAs)
```{r}
daily_steps_filled <- tapply(filled_data$steps,
                             filled_data$date,
                             function(x){sum(x)})
hist(daily_steps_filled, main = 'Steps Per Day Histogram', 
     xlab = '# steps per day')
```

### Mean of steps per day (filled NAs)
```{r}
mean(daily_steps_filled)
```

### Median of steps per day (filled NAs)
```{r}
median(daily_steps_filled)
```
### Do these values differ from the estimates from the first part of the assignment?


### What is the impact of imputing missing data on the estimates of the total daily number of steps?

#### Mean - Mean(filled NAs)
```{r}
mean(daily_steps) - mean(daily_steps_filled)
```
#### Median - Median(filled NAs)
```{r}
median(daily_steps) - median(daily_steps_filled)
```
## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lubridate)
is_weekend <- wday(as.Date(data$date))>5
data$is_weekend <- factor(is_weekend, labels = c('weekday', 'weekend'))
```

```{r}
interval_steps_week <- tapply(data$steps,
                              list(data$interval, data$is_weekend),
                              function(x){mean(x, na.rm = TRUE)})
par(mfrow=c(1,2))
plot(interval_steps_week[,'weekday'], 
     type = 'l',
     xlab = 'Interval', 
     ylab = 'Avarage # of steps',
     main='Weekday')
plot(interval_steps_week[,'weekend'], 
     type = 'l',
     xlab = 'Interval', 
     ylab = 'Avarage # of steps',
     main='Weekend')
```